### 2017-03-09 intro.tex
Finished intro

\section{The Importance of Feedback}\label{intro:fb}
\begin{figure}
    \includegraphics[width=\textwidth]{M82.eps}
	\caption[Massive outflows in M82]{The evidence for feedback is no clearer
	than here in the Cigar Galaxy, M82.  The Hubble Space Telescope reveals massive
	outflows of hot ionized gas through the red $H\alpha$ emission as gas is
	blasted out of the galaxy in a bipolar flow. \textit{Image credit: NASA, ESA
	and the Hubble Heritage Team STScI/AURA}}
\end{figure}
Evidence has been building for decades that galaxy and star formation is not a
one way process of gravitational collapse.  A number of internal processes
(feedback) within galaxies can liberate energy greater than the gravitational
binding energy of molecular clouds (and indeed even comparable to the binding
energy of the entire halo).  Analytic calculations that omit feedback processes
found that gas should rapidly collapse into the smallest halos at high redshift,
overproducing stars in small galaxies at high redshift (the ``cooling
catastrophe'') \citep{Cole2001,Benson2003}.  
Simulated galaxies develop unrealistically compact morphologies without feedback
\citep{Stinson2006}.  Without feedback, simulations of individual molecular
clouds have star formation rates an order of magnitude too high
\citep{Agertz2013}.  Aside from regulating star formation and the growth of
galaxies, there are also observations of the large-scale impact of feedback.
Lyman $\alpha$ forest observations probing the intergalactic medium (IGM) have
found that the IGM is polluted with metal ions at great distances from any
galaxies \citep{Sargent1988,Songaila1996,Dave1998}.  These metals must have been
ejected by galactic outflows from the ISM of galaxies where they formed. Figure
1.3 shows a composite image of the dwarf starburst M82, where fast, massive
outflows have been seen leaving the ISM.  These outflows have been observed as a
ubiquitous feature of starburst galaxies and quasar host galaxies
\citep{Veilleux2005,Werk2014}.

The two primary sources of feedback energy are massive stars and SMBHs.  Massive
stars ($M>5M_\odot$) experience short, violent lives.  Prior to their death,
they will drive fast $v>1000\;\rm{km/s}$ stellar winds \citep{Weaver1977}, and
ionize their surrounding gas with UV radiation \citep{Krumholz2009}.  At the end
of their lives, $4-30\;\rm{Myr}$ later, these stars die in energetic
core-collapse SN, liberating $\sim10^{51}\;\rm{erg}$
\citep{Wilson1985,Bethe1990}.  Figure 1.4 shows the total budget of available
energy liberated by a population of stars with a standard \citet{Chabrier2003}
IMF.  The total energy released by a population of stars with solar metallicity is
$\sim 2\times10^{51}\;\rm{erg/M_\odot}$.  Most of this energy is released as
photons, over the lifetime of each star.  These will couple poorly to the
surrounding ISM, and so have little impact on disrupting all but the densest
gas.  While stellar winds typically have comparable luminosity to supernovae,
they occur for an order of magnitude less time ($\sim4$ Myr vs. $\sim40$ Myr).
This means that supernovae, despite only contributing $\sim1\%$ of the total
energy released by a stellar population, may have the greatest impact on
galactic scales.  This energy imparts heat and momentum to the surrounding ISM,
and can accelerate cosmic rays in the shocks that result \citep{Bell1978}.  In
galaxies with actively growing SMBHs, the accretion disc that forms as gas
funnels into the black hole can reach temperatures exceeding $10^9\;\rm{K}$
through viscous heating \citep{Antonucci1993}.  This makes these accretion discs
some of the most luminous objects in the universe, and can heat the gas that
surrounds the galaxy to choke off the accretion of gas onto the galaxy, shutting
off star formation.

@@

On top of releasing just $10\%$ of the energy released by supernovae, stellar
winds are active during the early stages of a massive star's life.  This means
that these winds are released into the dense natal gas that massive star
clusters form in.  Stellar wind shocks within this dense molecular gas cool
efficiently, leading to much of the stellar wind energy being lost to radiation
that escapes the ISM completely.  Simulations of molecular clouds with feedback
from star clusters \citep{Gendelev2012,Rogers2013} have found that most stellar
wind energy is expended disrupting the molecular clouds the clusters live in.
While these winds (along with radiation pressure and UV-heated gas) do not
significantly escape the cluster, the channels they open within the cloud allow
a large fraction of SN energy to escape the dense molecular environment, while
the energy from winds and radiation are expended in disrupting the cloud.

@@

\subsection{Numerical Challenges for Hot Feedback}
The greatest difficulty with including the effects of feedback in simulations of
galaxy evolution is the combination of extreme gradients of temperature,
density, and velocity all in a relatively small mass/volume.  Without
overwhelmingly high resolution, hot feedback gas will be numerically mixed with
cold ISM, often resulting in gas with temperatures of $\sim10^5\;\rm{K}$, at the
peak of the ISM cooling curve where cooling times are extremely short.  This
results in feedback energy being completely lost due to radiative cooling, as
was first seen by \citet{Katz1992}.  Indeed, as \citet{Thacker2000} showed, how
feedback energy is deposited into the ISM has a huge impact on whether that
feedback has any effect.  This problem has been tackled using a few different
strategies.

Since overcooling is the problem, many models have attacked this directly by
either temporarily disabling cooling \citep{Thacker2000,Stinson2006} or
depositing feedback energy into a second, nonthermal energy component
\citep{Agertz2013}.  While this does limit radiative losses, it produces gas
that exists in an unphysical region of the $\rho-T$ phase diagram, which may not
have the correct entropy to escape buoyantly.  It also has the potential to
introduce an \textit{undercooling} problem as resolution becomes higher and
feedback-heated gas becomes resolvable.  Alternatively, energy may be injected
not as heat, but as momentum, which naturally does not cool
\citep{Navarro1993,Mihos1994,Scannapieco2006,DallaVecchia2008,Dubois2008}.
Unfortunately, this adds the additional complexity of momentum cancellation (how
to treat neighbouring stars/clusters, where two stars are depositing momentum
with opposing directions), and if this momentum results in strong shocks, these
shocks will convert the momentum right back into thermal energy that can radiate
away the feedback energy the model is designed to preserve \citep{Durier2012}.
This often means kinetic/momentum feedback models must be combined with
hydrodynamic decoupling that renders feedback-heated gas into a form that does
not interact with the surrounding ISM \citep{Springel2003,Vogelsberger2013}.
Naturally, this makes outflow properties strongly influenced by numerical
choices \citep{DallaVecchia2008}.  A third choice is to stochastically deposit
feedback energy only when enough energy is available to produce gas temperatures
high enough for cooling times to be short \citep{DallaVecchia2012,Crain2015}.
In these models, SN events are effectively grouped together, somewhat decoupling
feedback spatially from star formation.  The choice of temperature to heat gas
to is a purely numerical parameter, and can dramatically change the
effectiveness of these models.  Finally, some \citep{Springel2003,Murante2015}
attempt to directly address the unphysical numerical mixing of hot and cold gas
by separating the two phases and tracking each component separately.  This
completely solves the overcooling problem, and allows the two components to cool
radiatively.   Unfortunately, it also means that cold star-forming gas is
permanently coupled to hot feedback-heated gas: preventing the simulation of  a
resolved multiphase ISM, and effectively making the cold/warm ISM an anchor
holding down potential hot outflows.  This means these multiphase models have to
be coupled with parameterized models to capture outflows \citep{Springel2003}.
Often, these simple models simply choose a fraction of feedback energy to be
deposited as kinetic ``kicks'' to a some gas particles.  Like other kinetic
feedback models, this is frequently combined with hydrodynamical decouplings
\citep{Vogelsberger2013}.

All of these models attempt to solve the problem of overcooling with unphysical
approximations: disabled cooling, hydrodynamic decoupling, stochastic deposition
of feedback energy, and imposed multiphase ISM models.  These choices make it
difficult to understand what details of simulated galaxy evolution are a result
of these numerical free parameters, as opposed to the actual physics of
self-regulated star formation and SN-driven outflows.  What is needed is a model
that captures the effects of feedback while matching as closely as possible the
actual physical processes involved below the resolution scale.  Simulations that
rely on numerically-motivated, complex subgrid models are devoid of serious
predictive power.  When the results obtained depend heavily of the tuning of
multiple free parameters, those results can at best probe one point in the space
of the possible, rather than hew away at that space to discover the region of
the probable, and ultimately, the truth.  As Von Neumann famously quipped to
Fermi, ``With four parameters I can fit an elephant, and with five I can make
him wiggle his trunk.''

@@

The structure and evolution of superbubbles have been the subject of a great
deal of analytic study.  \citet{Castor1975} found that the time evolution of
superbubbles has a similarity solution, the \citet{Sedov1959} adiabatic blast
wave solution: 
\begin{equation}
    R(t)=27n_0^{-1/5}L_{36}^{1/5}t_6^{3/5}\;\rm{pc} 
\end{equation}
\begin{equation}
    V(t)=16n_0^{-1/5}L_{36}^{1/5}t_6^{-2/5}\;\rm{km/s}
\end{equation}
$n_0$ is the ambient density in $\rm{cm^{-3}}$, $L_{36}$ is the luminosity of
the star cluster in $10^{36}\;\rm{erg/s}$, and $t_6$ is the time in
$10^6\;\rm{yr}$.  The evolution of hot mass in the bubble, and the effects of
cooling in the swept-up shell, were examined in \citet{Weaver1977}.  A major
result of this study was the realization that thermal conduction between the hot
bubble and cold shell will result in an evaporative mass flux, pulling material
into the hot bubble.  This evaporated mass will quickly dominate over the small
amount of ejecta from stellar winds (and later SN).  The \citet{Weaver1977}
results were extended to clustered SN by \citet{MacLow1988}.  They calculated
that the typical cooling time for the hot bubble was $\sim15\;\rm{Myr}$, with a
weak dependence on the total luminosity and ambient density.  Before significant
cooling can occur in the hot interior, $\sim35\%$ of the total feedback
luminosity is lost to cooling of the swept-up shell, which cools in
$\sim10^4\;\rm{yr}$.  This means that, rather than the $1-10\%$ of energy 
preserved from a single supernova blast \citep{Chevalier1974}, $\sim65\%$ of the
initial energy budget is retained in the kinetic energy of the cold shell and
thermal energy of the hot interior.  \citet{MacLow1988} also demonstrated that,
as superbubbles do not have surface tension, they are not depressurized by cold
clouds, and instead simply deform around them while continuing to expand.
\citet{Silich1996} studied the effects of an ISM permeated with cold clumps in
greater detail, finding that the general behaviour discussed in
\citet{MacLow1988} held even in the case of an inhomogeneous ISM.  Even with
$50\%$ of the ISM mass trapped in cold clouds (which penetrate the cold shell of
the superbubble), the total hot mass within the superbubble will only vary by
$\sim20\%$.

@@

\citet{DallaVecchia2012} showed that the temperature of feedback-heated gas can
ultimately determine its effectiveness in driving outflows and regulating star
formation, even with constant feedback energy.  The self-regulating temperature
of superbubbles provides the natural mechanism for setting this temperature:
thermal evaporation.

>>>>

\section{The Importance of Feedback}\label{intro:fb}
\begin{figure}
    \includegraphics[width=\textwidth]{M82.eps}
	\caption[Massive outflows in M82]{The evidence for feedback is no clearer
	than here in the Cigar Galaxy, M82.  The Hubble Space Telescope reveals massive
	outflows of hot ionized gas through the red $H\alpha$ emission as gas is
	blasted out of the galaxy in a bipolar flow. \textit{Image credit: NASA, ESA
	and the Hubble Heritage Team STScI/AURA}}
\end{figure}
Evidence has been building for decades that galaxy and star formation is not a
one way process of gravitational collapse.  A number of internal processes can
release a tremendous amount of energy back into the ISM of the galaxy, stirring
or even ejecting that gas.  Analytic calculations that omit feedback processes
found that gas should rapidly collapse into the smallest halos that can cool
radiatively (halos with a virial temperature of $\sim 10^4\;K$, or a mass of
$\sim10^8\Msun$).  This leads to an overproduction of stars in small galaxies at
high redshift (the ``cooling catastrophe'') \citep{Cole2001,Benson2003}.  The
observational signatures of feedback are abundant.  Lyman $\alpha$ forest
observations probing the intergalactic medium (IGM) have found that the IGM is
polluted with metal ions at great distances from any galaxies
\citep{Sargent1988,Songaila1996,Dave1998}.  These metals must have been ejected
by galactic outflows from the ISM of galaxies where they formed. Figure 1.3
shows a composite image of the dwarf starburst M82, where fast, massive outflows
have been seen leaving the ISM.  These outflows have been observed as a
ubiquitous feature of starburst galaxies and quasar host galaxies
\citep{Veilleux2005,Werk2014}.  Numerical models also have pointed towards
feedback as an essential ingrediant in forming both realistic galaxies and the
structures within them. Simulated galaxies develop unrealistically compact
morphologies without feedback \citep{Stinson2006}.  Inside of the galaxy,
simulations of individual molecular clouds have star formation rates an order of
magnitude too high \citep{Agertz2013} when feedback is omitted.  Clearly,
if we wish to accurately simulate galaxies, we must also accurately model the
feedback processes that shape those galaxies.

The two primary sources of feedback energy are massive stars and SMBHs.  Massive
stars ($M>5M_\odot$) experience short, violent lives.  Prior to their death,
they will drive fast $v>1000\;\rm{km/s}$ stellar winds \citep{Weaver1977}, and
ionize their surrounding gas with UV radiation \citep{Krumholz2009}.  At the end
of their lives, $4-30\;\rm{Myr}$ later, these stars die in energetic
core-collapse SN, liberating $\sim10^{51}\;\rm{erg}$
\citep{Wilson1985,Bethe1990}.  Figure 1.4 shows the total budget of available
energy liberated by a population of stars with a standard \citet{Chabrier2003}
IMF.  The total energy released by a population of stars with solar metallicity is
$\sim 2\times10^{51}\;\rm{erg/M_\odot}$.  Most of this energy is released as
photons, over the lifetime of each star.  These will both couple weakly with the
ISM, depositing little energy, and only heat gas to a relatively modest
temperature, much cooler than the virial temperature for galaxies like our own.
While stellar winds from a cluster will have comparable luminosity to
supernovae, two factors limit the effectiveness of these winds at driving
galaxy-scale outflows.  First, because these stellar winds are produced early,
they couple to the dense environment of the cluster's natal cloud, resulting in
orders of magnitude shorter cooling times for the hot bubbles they drive.
Second, the vast majority of the energy liberated in stellar winds comes from
the most massive stars, that die within $4\Myr$, giving stellar winds a total
energy budget of $10\%$ of the supernovae budget \citet{Leitherer1999}.
This means that supernovae, despite only contributing $\sim1\%$ of the total
energy released by a stellar population, may have the greatest impact on
galactic scales.  This energy imparts heat and momentum to the surrounding ISM,
and can accelerate cosmic rays in the shocks that result \citep{Bell1978}.  In
galaxies with actively growing SMBHs, the accretion disc that forms as gas
funnels into the black hole can reach temperatures exceeding $10^9\;\rm{K}$
through viscous heating \citep{Antonucci1993}.  This makes these accretion discs
some of the most luminous objects in the universe, and can heat the gas that
surrounds the galaxy to choke off the accretion of gas onto the galaxy, shutting
off star formation.

@@

On top of releasing just $10\%$ of the energy released by supernovae, stellar
winds are active during the early stages of a massive star's life.   Simulations
of molecular clouds with feedback from star clusters
\citep{Gendelev2012,Rogers2013} have found that most stellar wind energy is
expended disrupting the molecular clouds the clusters live in.  While these
winds (along with radiation pressure and UV-heated gas) do not significantly
escape the cluster, the channels they open within the cloud allow a large
fraction of SN energy to escape the dense molecular environment, while the
energy from winds and radiation are expended in disrupting the cloud.  Thus, for
galaxy-scale simulations that do not resolve the internal structure of
individual molecular clouds, the effects of UV radiation and stellar winds can
be safely omitted.

@@

\subsection{Numerical Challenges for Hot Feedback}
The greatest difficulty with including the effects of feedback in simulations of
galaxy evolution is the combination of extreme gradients of temperature,
density, and velocity all in a relatively small mass/volume.  Without
overwhelmingly high resolution, hot feedback gas will be numerically mixed with
cold ISM, often resulting in gas with relatively high density and temperatures
of $\sim10^5\;\rm{K}$, where radiative cooling times are extremely short.  This
results in feedback energy being completely lost due to radiative cooling, as
was first seen by \citet{Katz1992}.  Indeed, as \citet{Thacker2000} showed, how
feedback energy is deposited into the ISM has a huge impact on whether that
feedback has any effect.  This problem has been tackled using a few different
strategies.

Since overcooling is the problem, many models have attacked this directly by
either temporarily disabling cooling \citep{Thacker2000,Stinson2006} or
depositing feedback energy into a second, nonthermal energy component
\citep{Agertz2013}.  While this does limit radiative losses, it produces gas
that exists in an unphysical region of the $\rho-T$ phase diagram.  This gas may
lack the high temperatures and velocities required to actually escape the disc.
It also has the potential to introduce an \textit{undercooling} problem as
resolution becomes higher and feedback-heated gas becomes resolvable.
Alternatively, energy may be injected not as heat, but as momentum, which
naturally does not cool
\citep{Navarro1993,Mihos1994,Scannapieco2006,DallaVecchia2008,Dubois2008}.
Unfortunately, this adds the additional complexity of momentum cancellation (how
to treat neighbouring stars/clusters, where two stars are depositing momentum
with opposing directions), and if this momentum results in strong shocks, these
shocks will convert the momentum back into thermal energy that can radiate
away the very energy the model is designed to preserve \citep{Durier2012}.
This often means kinetic/momentum feedback models must be combined with
hydrodynamic decoupling that renders feedback-heated gas into a form that does
not interact with the surrounding ISM \citep{Springel2003,Vogelsberger2013}.
Naturally, this makes outflow properties strongly influenced by numerical
choices \citep{DallaVecchia2008}.  A third choice is to stochastically deposit
feedback energy only when enough energy is available to produce gas temperatures
high enough for cooling times to be short \citep{DallaVecchia2012,Crain2015}.
In these models, SN events are effectively grouped together, somewhat decoupling
feedback spatially from star formation.  The choice of temperature to heat gas
to is a purely numerical parameter, and can dramatically change the
effectiveness of these models.  Finally, some \citep{Springel2003,Murante2015}
attempt to directly address the unphysical numerical mixing of hot and cold gas
by separating the two phases and tracking each component separately.  This
completely solves the overcooling problem, and allows the two components to cool
radiatively.   Unfortunately, it also means that cold star-forming gas is
permanently coupled to hot feedback-heated gas: preventing the simulation of  a
resolved multiphase ISM, and effectively making the cold/warm ISM an anchor
holding down potential hot outflows.  This means these multiphase models have to
be coupled with parameterized models to capture outflows \citep{Springel2003}.
Often, these simple models simply choose a fraction of feedback energy to be
deposited as kinetic ``kicks'' to a some gas particles.  Like other kinetic
feedback models, this is frequently combined with hydrodynamical decouplings
\citep{Vogelsberger2013}.

All of these models attempt to solve the problem of overcooling with unphysical
approximations: disabled cooling, hydrodynamic decoupling, stochastic deposition
of feedback energy, and imposed multiphase ISM models.  These choices make it
difficult to understand what details of simulated galaxy evolution are a result
of these numerical free parameters, as opposed to the actual physics of
self-regulated star formation and SN-driven outflows.  What is needed is a model
that captures the effects of feedback while matching as closely as possible the
actual physical processes involved below the resolution scale.  Simulations that
rely on complex, numerically-motivated subgrid models are devoid of serious
predictive power.  When the results obtained depend heavily of the tuning of
multiple free parameters, those results can at best probe one point in the space
of the possible, rather than hew away at that space to discover the region of
the probable, and ultimately, the truth.

@@

The structure and evolution of superbubbles have been the subject of a great
deal of analytic study.  \citet{Castor1975} found that the time evolution of
superbubbles has a similarity solution, similar to the \citet{Sedov1959}
adiabatic blast wave solution: 
\begin{equation}
    R(t)=27n_0^{-1/5}L_{36}^{1/5}t_6^{3/5}\;\rm{pc} 
\end{equation}
\begin{equation}
    V(t)=16n_0^{-1/5}L_{36}^{1/5}t_6^{-2/5}\;\rm{km/s}
\end{equation}
$n_0$ is the ambient density in $\rm{cm^{-3}}$, $L_{36}$ is the luminosity of
the star cluster in $10^{36}\;\rm{erg/s}$, and $t_6$ is the time in
$10^6\;\rm{yr}$.  The evolution of hot mass in the bubble, and the effects of
cooling in the swept-up shell, were examined in \citet{Weaver1977}.  The
\citet{Weaver1977} results were extended to clustered SN by \citet{MacLow1988}.
They calculated that the typical cooling time for the hot bubble was
$\sim15\;\rm{Myr}$, with a weak dependence on the total luminosity and ambient
density.  Before significant cooling can occur in the hot interior, $\sim35\%$
of the total feedback luminosity is lost to cooling of the swept-up shell, which
cools in $\sim10^4\;\rm{yr}$.  This means that, rather than the $1-10\%$ of
energy preserved from a single supernova blast \citep{Chevalier1974}, $\sim65\%$
of the initial energy budget is retained in the kinetic energy of the cold shell
and thermal energy of the hot interior.  \citet{MacLow1988} also demonstrated
that, as superbubbles do not have surface tension, they are not depressurized by
cold clouds, and instead simply deform around them while continuing to expand.
\citet{Silich1996} studied the effects of an ISM permeated with cold clumps in
greater detail, finding that the general behaviour discussed in
\citet{MacLow1988} held even in the case of an inhomogeneous ISM.  Even with
$50\%$ of the ISM mass trapped in cold clouds (which penetrate the cold shell of
the superbubble), the total hot mass within the superbubble will only vary by
$\sim20\%$.

@@

\citet{DallaVecchia2012} showed that the temperature of feedback-heated gas can
ultimately determine its effectiveness in driving outflows and regulating star
formation, even with constant feedback energy.  The self-regulating temperature
of superbubbles provides the natural mechanism for setting this temperature:
thermal evaporation.  By including the effects of thermal evaporation between
unresolved hot bubbles and the surrounding cool ISM, feedback energy can be
deposited into the correct amount of ISM mass, giving realistic temperatures,
densities, and cooling times for feedback-heated gas.  This superbubble feedback
model is described in detail, and applied to the problem of Milky Way-like
galaxy formation in the chapters that follow.
